---
title: "Bayesian_Modeling_Project_Code"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Read in packages
```{r setup, include=FALSE}
library(brms)
library(rstan)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
library(dplyr)
library(fpp2)
#library(bayesforecast)
library(ggthemes)
library(bayesplot)
library(here)


options(max.print = 100000)
```

# clear environment
```{r}
rm(list = ls(all.names = TRUE))
```


# Read in data
```{r}

# read in from github
url <- "https://raw.githubusercontent.com/joshrrosen/Bayesian_Modeling_Project/main/Bikeshare_data.csv"
bikeshare_data <- read.csv(url)


```


# Create individual dataframes
```{r}

# boston
df_boston <- bikeshare_data %>%
  filter(CITY == "BOSTON")


## --------------------------------------------------

# cambridge
df_cambridge <- bikeshare_data %>%
  filter(CITY == "CAMBRIDGE") 


## --------------------------------------------------

# Brookline 
df_brookline <- bikeshare_data %>%
  filter(CITY == "BROOKLINE") 


## --------------------------------------------------

# somerville 
df_somerville <- bikeshare_data %>%
  filter(CITY == "SOMERVILLE") 

```



# updated dataframe
```{r}

#detach("package:bayesforecast", unload = TRUE)

# create a dataframe 
merged_df <- rbind(df_boston, df_cambridge, df_brookline, df_somerville)

merged_df <- merged_df %>%
  mutate(city_id = as.numeric(as.factor(CITY))) %>%
  filter(After == 0) # only pre-bike share period


```



# predict cambridge 
```{r}

# isolate cambridge data
df_cambridge1 <- merged_df %>%
  filter(CITY == "CAMBRIDGE")

# get priors
brms::get_prior(Total_Injuries ~ 0 + arma(p = 1, q = 1), data = df_cambridge1)


stan_dat2 <- list(y = df_cambridge1$Total_Injuries, 
                 T = length(df_cambridge1$Total_Injuries),
                 n_new = 10) # everything over 30 is a prediction

fit3 <- stan(file = "C:\\Users\\User\\Desktop\\Bayesian Stats\\Bayesian_Modeling_Project\\Bayesian_Modeling_Project\\single_city_prediction_cambridge.stan",
             iter = 2000,
             chains = 4,
             data = stan_dat2)


#summary(fit3, pars = c("nu", "predict"), probs = c(.1, .9))

# create dataframe of predictions
cambridge_predict <- as.data.frame(summary(fit3, pars = c("predict"), probs = c(.1, .9)))


cambridge_predict <- cambridge_predict %>%
  mutate(Count = c(1:40))


# add true injured to file
cambridge_predict <- cambridge_predict %>%
  mutate(true_injured = df_cambridge$Total_Injuries[1:40])


# plot predictions with confidence interval
plot(cambridge_predict$Count, cambridge_predict$summary.mean, type = "b", col = "blue", ylim = c(18, 70), lty = 1, lwd = 1, 
      xlab = "Months",
      ylab = "Traffic Injuries")
lines(cambridge_predict$Count, cambridge_predict$summary.10., col = "red")
lines(cambridge_predict$Count, cambridge_predict$summary.90., col = "red")
abline(v=31, lty=2, lwd=3)
legend("topright", legend = c("10 pct", "Mean Prediction", "90 pct"),
       col = c("red", "blue"), lty = 1:2, cex = 0.8)
title(main = "Injury Prediction: Cambridge")


# plot comparison 
plot(cambridge_predict$Count, cambridge_predict$summary.mean, type = "b", col = "blue", ylim = c(18, 70), lty = 1, lwd = 1, 
      xlab = "Months",
      ylab = "Traffic Injuries")
lines(cambridge_predict$Count, cambridge_predict$true_injured, type = "b", col = "red", lty = 1, lwd = 1)
abline(v=31, lty=2, lwd=3)
legend("topright", legend = c("Actual", "Predicted"),
       col = c("red", "blue"), lty = 1:2, cex = 0.8)
title(main = "Injury Prediction: Cambridge")


# calculate the mean absolute percentage error
mape <- mean(abs((cambridge_predict$summary.mean - cambridge_predict$true_injured) / cambridge_predict$true_injured))
mape

# means absolute error before 
mae <- mean(abs(cambridge_predict$summary.mean[1:30] - cambridge_predict$true_injured[1:30]))
mae

#After 
mae <- mean(abs(cambridge_predict$summary.mean[31:40] - cambridge_predict$true_injured[31:40]))
mae

```

# Boston predict
```{r}

# isolate city data
df_boston1 <- merged_df %>%
  filter(CITY == "BOSTON")

# get priors
brms::get_prior(Total_Injuries ~ 0 + arma(p = 1, q = 1), data = df_boston1)

stan_dat_bos <- list(y = df_boston1$Total_Injuries, 
                 T = length(df_boston1$Total_Injuries),
                 n_new = 10)

fit_bos <- stan(file = "C:\\Users\\User\\Desktop\\Bayesian Stats\\Bayesian_Modeling_Project\\Bayesian_Modeling_Project\\single_city_prediction_boston.stan",
             iter = 2000,
             chains = 4,
             data = stan_dat_bos)


#summary(fit_bos, pars = c("nu", "predict"), probs = c(.1, .9))


# create dataframe of predictions
boston_predict <- as.data.frame(summary(fit_bos, pars = c("predict"), probs = c(.1, .9)))


boston_predict <- boston_predict %>%
  mutate(Count = c(1:28))


# add true injured to file
boston_predict <- boston_predict %>%
  mutate(true_injured = df_boston$Total_Injuries[1:28])


# plot predictions with confidence interval
plot(boston_predict$Count, boston_predict$summary.mean, type = "b", col = "blue", ylim = c(0,320), lty = 1, lwd = 1, 
      xlab = "Months",
      ylab = "Traffic Injuries")
lines(boston_predict$Count, boston_predict$summary.10., col = "red")
lines(boston_predict$Count, boston_predict$summary.90., col = "red")
abline(v=19, lty=2, lwd=3)
legend("topright", legend = c("10 pct", "Mean Prediction", "90 pct"),
       col = c("red", "blue"), lty = 1:2, cex = 0.8)
title(main = "Injury Prediction: Boston")


# plot comparison 
plot(boston_predict$Count, boston_predict$summary.mean, type = "b", col = "blue", ylim = c(0,320), lty = 1, lwd = 1, 
      xlab = "Months",
      ylab = "Traffic Injuries")
lines(boston_predict$Count, boston_predict$true_injured, type = "b", col = "red", lty = 1, lwd = 1)
abline(v=19, lty=2, lwd=3)
legend("topright", legend = c("Actual", "Predicted"),
       col = c("red", "blue"), lty = 1:2, cex = 0.8)
title(main = "Injury Prediction: Boston")

mape <- mean(abs((boston_predict$summary.mean - boston_predict$true_injured) / boston_predict$true_injured))
mape


# means absolute error before 
mae <- mean(abs(boston_predict$summary.mean[1:18] - boston_predict$true_injured[1:18]))
mae

#After 
mae <- mean(abs(boston_predict$summary.mean[19:28] - boston_predict$true_injured[19:28]))
mae
```


# predict somerville
```{r}

# isolate city data
df_somerville1 <- merged_df %>%
  filter(CITY == "SOMERVILLE")

# get priors
brms::get_prior(Total_Injuries ~ 0 + arma(p = 1, q = 1), data = df_somerville1)

stan_dat_som <- list(y = df_somerville1$Total_Injuries, 
                 T = length(df_somerville1$Total_Injuries),
                 n_new = 10) 

fit_som <- stan(file = "C:\\Users\\User\\Desktop\\Bayesian Stats\\Bayesian_Modeling_Project\\Bayesian_Modeling_Project\\single_city_prediction_somerville.stan",
             iter = 2000,
             chains = 4,
             data = stan_dat_som)


#summary(fit_som, pars = c("nu", "predict"), probs = c(.1, .9))


# create dataframe of predictions
somerville_predict <- as.data.frame(summary(fit_som, pars = c("predict"), probs = c(.1, .9)))


somerville_predict <- somerville_predict %>%
  mutate(Count = c(1:41))


# add true injured to file
somerville_predict <- somerville_predict %>%
  mutate(true_injured = df_somerville$Total_Injuries[1:41])


# plot predictions with confidence interval
plot(somerville_predict$Count, somerville_predict$summary.mean, type = "b", col = "blue", ylim = c(0,50), lty = 1, lwd = 1, 
      xlab = "Months",
      ylab = "Traffic Injuries")
lines(somerville_predict$Count, somerville_predict$summary.10., col = "red")
lines(somerville_predict$Count, somerville_predict$summary.90., col = "red")
abline(v=32, lty=2, lwd=3)
legend("topright", legend = c("10 pct", "Mean Prediction", "90 pct"),
       col = c("red", "blue"), lty = 1:2, cex = 0.8)
title(main = "Injury Prediction: Somerville")


# plot comparison 
plot(somerville_predict$Count, somerville_predict$summary.mean, type = "b", col = "blue", ylim = c(0,50), lty = 1, lwd = 1, 
      xlab = "Months",
      ylab = "Traffic Injuries")
lines(somerville_predict$Count, somerville_predict$true_injured, type = "b", col = "red", lty = 1, lwd = 1)
abline(v=32, lty=2, lwd=3)
legend("topright", legend = c("Actual", "Predicted"),
       col = c("red", "blue"), lty = 1:2, cex = 0.8)
title(main = "Injury Prediction: Somerville")

mape <- mean(abs((somerville$summary.mean - somerville$true_injured) / somerville$true_injured))
mape


# means absolute error before 
mae <- mean(abs(somerville_predict$summary.mean[1:31] - somerville_predict$true_injured[1:31]))
mae

#After 
mae <- mean(abs(somerville_predict$summary.mean[32:41] - somerville_predict$true_injured[32:41]))
mae

```


# predict brookline
```{r}

# isolate city data
df_brookline1 <- merged_df %>%
  filter(CITY == "BROOKLINE")

# get priors
brms::get_prior(Total_Injuries ~ 0 + arma(p = 1, q = 1), data = df_brookline1)

stan_dat_brook <- list(y = df_brookline1$Total_Injuries, 
                 T = length(df_brookline1$Total_Injuries),
                 n_new = 10) 

fit_brook <- stan(file = "C:\\Users\\User\\Desktop\\Bayesian Stats\\Bayesian_Modeling_Project\\Bayesian_Modeling_Project\\single_city_prediction_brookline.stan",
             iter = 1000,
             chains = 4,
             data = stan_dat_brook)


#summary(fit_brook, pars = c("nu", "predict"), probs = c(.1, .9))


# create dataframe of predictions
brookline_predict <- as.data.frame(summary(fit_brook, pars = c("predict"), probs = c(.1, .9)))


brookline_predict <- brookline_predict %>%
  mutate(Count = c(1:41))


# add true injured to file
brookline_predict <- brookline_predict %>%
  mutate(true_injured = df_brookline$Total_Injuries[1:41])


# plot predictions with confidence interval
plot(brookline_predict$Count, brookline_predict$summary.mean, type = "b", col = "blue", ylim = c(0,30), lty = 1, lwd = 1, 
      xlab = "Months",
      ylab = "Traffic Injuries")
lines(brookline_predict$Count, brookline_predict$summary.10., col = "red")
lines(brookline_predict$Count, brookline_predict$summary.90., col = "red")
abline(v=32, lty=2, lwd=3)
legend("topright", legend = c("10 pct", "Mean Prediction", "90 pct"),
       col = c("red", "blue"), lty = 1:2, cex = 0.8)
title(main = "Injury Prediction: Brookline")


# plot comparison 
plot(brookline_predict$Count, brookline_predict$summary.mean, type = "b", col = "blue", ylim = c(0,30), lty = 1, lwd = 1, 
      xlab = "Months",
      ylab = "Traffic Injuries")
lines(brookline_predict$Count, brookline_predict$true_injured, type = "b", col = "red", lty = 1, lwd = 1)
abline(v=32, lty=2, lwd=3)
legend("topright", legend = c("Actual", "Predicted"),
       col = c("red", "blue"), lty = 1:2, cex = 0.8)
title(main = "Injury Prediction: Brookline")


mape <- mean(abs((brookline_predict$summary.mean - brookline_predict$true_injured) / brookline_predict$true_injured))
mape


# means absolute error before 
mae <- mean(abs(brookline_predict$summary.mean[1:30] - brookline_predict$true_injured[1:30]))
mae

#After 
mae <- mean(abs(brookline_predict$summary.mean[31:40] - brookline_predict$true_injured[31:40]))
mae

```


# plots for just pre-intervention trends
```{r}
library(bayesforecast)

# boston

# diagnostics
ggacf(df_boston1$Total_Injuries) # autocorrelation function
ggpacf((df_boston1$Total_Injuries)) # partial autocorrelation function

# time series data
bos_y <- ts(df_boston1$Total_Injuries, start=1, frequency=12)

# seasonal plot
ggseasonplot(bos_y, polar=TRUE) +
  ylab("Total Injuries") +
  ggtitle("Polar seasonal plot: Traffic Injuries")


## --------------------------------------------------

# cambridge

# diagnostics 
ggacf(df_cambridge1$Total_Injuries) # autocorrelation function
ggpacf((df_cambridge1$Total_Injuries)) # partial autocorrelation function

# time series
cam_y <- ts(df_cambridge1$Total_Injuries, start=1, frequency=12)

# seasonal
ggseasonplot(cam_y, polar=TRUE) +
  ylab("Total Injuries") +
  ggtitle("Polar seasonal plot: Traffic Injuries")

## --------------------------------------------------

# brookline

ggacf(df_brookline1$Total_Injuries) # autocorrelation function
ggpacf((df_brookline1$Total_Injuries)) # partial autocorrelation function

# time series
bro_y <- ts(df_brookline1$Total_Injuries, start=1, frequency=12)

# seasonal
ggseasonplot(bro_y, polar=TRUE) +
  ylab("Total Injuries") +
  ggtitle("Polar seasonal plot: Traffic Injuries")

## --------------------------------------------------

# somerville 

ggacf(df_somerville1$Total_Injuries) # autocorrelation function
ggpacf((df_somerville1$Total_Injuries)) # partial autocorrelation function

# time series
som_y <- ts(df_somerville1$Total_Injuries, start=1, frequency=12)

# seasonal
ggseasonplot(som_y, polar=TRUE) +
  ylab("Total Injuries") +
  ggtitle("Polar seasonal plot: Traffic Injuries") 


```


# Single Plot
```{r}


# boston
df_boston2 <- df_boston1 %>%
  mutate(months = seq(1, 18, by = 1)) %>%
  mutate(bos_injured = Total_Injuries) %>%
  select(months, bos_injured)

# brookline
df_brookline2 <- df_brookline1%>%
  mutate(months = seq(1, 31, by =1)) %>%
  mutate(brook_injured = Total_Injuries) %>%
  select(months, brook_injured)

df_somerville2 <- df_somerville1 %>%
  mutate(months = seq(1, 31, by =1)) %>%
  mutate(som_injured = Total_Injuries) %>%
  select(months, som_injured) 

df_cambridge2 <- df_cambridge1 %>%
  mutate(months = seq(1,30, by = 1)) %>%
  mutate(cam_injured = Total_Injuries) %>%
  select(months, cam_injured)

bos_brook_df <- merge(df_boston2, df_brookline2, by = "months", all = TRUE)

bos_brook_cam_df <- merge(bos_brook_df, df_cambridge2, by = "months", all = TRUE)

dated_df <- merge(bos_brook_cam_df, df_somerville2, by = "months", all = TRUE)



ggplot(dated_df, aes(months)) +
  geom_line(aes(y = bos_injured), color = "black") + 
  geom_line(aes(y = brook_injured), color = "red") +
  geom_line(aes(y = cam_injured), color = "blue") +
  geom_line(aes(y = som_injured), color = "green") +
  xlab("Months") + ylab("Injuries") +
  ggtitle("Traffic Injuries Prior to Treatment") +
  theme_few() + scale_colour_few()
  

```



# Modeling the autoregressive term hierarchically
```{r}

# create data
merged_df2 <- merged_df %>%
  filter(time_since_treatment >= -10)


stan_data_H <- list(y = as.array(merged_df2$Total_Injuries), 
                 T = 10,
                 N = 40) # 40


Hierarchical_fit1 <- stan(file = "C:\\Users\\User\\Desktop\\Bayesian Stats\\Bayesian_Modeling_Project\\Bayesian_Modeling_Project\\Hierarhical_mod_standard.stan",
             data = stan_data_H,
             chains = 4,
             iter = 2000)

summary(Hierarchical_fit1, pars = "y_pred", probs = c(.1, .9))


######## Predicting Out

stan_data_h2 <- list(y = as.array(merged_df2$Total_Injuries), 
                 T = 10,
                 N = 40,
                 M = 10)

Hierarchical_fit2 <- stan(file = "C:\\Users\\User\\Desktop\\Bayesian Stats\\Bayesian_Modeling_Project\\Bayesian_Modeling_Project\\Hierarchical_predictions.stan",
                 data = stan_data_h2,
                 chains = 4,
                 iter = 2000)

# only predicting for Somerville
summary(Hierarchical_fit2, pars = c("y_pred", "y_pred_new_final"), probs = c(.1, .9))

```


# Hierachical vs. ARMA(1,1) model (Boston)
```{r}

df <- as.data.frame(summary(Hierarchical_fit1, pars = "y_pred", probs = c(.1, .9)))

df <- df %>%
  mutate(Count = c(1:40)) %>%
  select(summary.mean, summary.90., summary.10., Count)
  
df <- df[-c(11:40),]

df <- df %>%
  mutate(true_injured = df_boston$Total_Injuries[9:18]) %>%
  mutate(other_mod = boston_predict$summary.mean[9:18]) %>%
  mutate(other_90 = boston_predict$summary.90.[9:18]) %>%
  mutate(other_10 = boston_predict$summary.10.[9:18]) 



# plot predictions with confidence interval
plot(df$Count, df$summary.mean, type = "b", col = "blue", ylim = c(75,300), lty = 1, lwd = 1, 
      xlab = "Months",
      ylab = "Traffic Injuries")
lines(df$Count, df$summary.10., col = "red")
lines(df$Count, df$summary.90., col = "red")
lines(df$Count, df$other_mod, col = "black", type = "b", lwd = 1)
lines(df$Count, df$other_10, col = 'black')
lines(df$Count, df$other_90, col = "black")
lines(df$Count, df$true_injured, col = "green", type = "b", lwd = 1)
legend("topright", legend = c("Hierarchical: 10 pct", "Hierarchical: Mean Prediction", "Hierarchical: 90 pct", "True Injured", "Original Model"),
       col = c("red", "blue", "red", "green", "black"), lty = 1:2, cex = 0.8)
title(main = "Injury Prediction: Boston")


# without intervals 
plot(df$Count, df$summary.mean, type = "b", col = "blue", ylim = c(75,300), lty = 1, lwd = 1, 
      xlab = "Months",
      ylab = "Traffic Injuries")
lines(df$Count, df$other_mod, col = "black", type = "b", lwd = 1)
lines(df$Count, df$true_injured, col = "green", type = "b", lwd = 1)
legend("topright", legend = c("Original Model", "Hierarchical Prediction", "True Injured"),
       col = c("black", "blue", "green"), lty = 1:2, cex = 0.8)
title(main = "Injury Prediction: Boston (No CIs)")

```

# Predict Boston and Compare
```{r}

df <- as.data.frame(summary(Hierarchical_fit2, pars = c("y_pred","y_pred_new_final"), probs = c(.1, .9)))

df <- df %>%
  mutate(Count = c(1:50)) %>%
  select(summary.mean, summary.90., summary.10., Count)
  
df <- df[-c(11:40),]

df <- df %>%
  mutate(true_injured = df_boston$Total_Injuries[9:28]) %>%
  mutate(other_mod = boston_predict$summary.mean[9:28]) %>%
  mutate(other_90 = boston_predict$summary.90.[9:28]) %>%
  mutate(other_10 = boston_predict$summary.10.[9:28]) 

df <- df %>%
  mutate(Count = c(1:20))

# plot predictions with confidence interval
plot(df$Count, df$summary.mean, type = "b", col = "blue", ylim = c(75,300), lty = 1, lwd = 1, 
      xlab = "Months",
      ylab = "Traffic Injuries")
lines(df$Count, df$summary.10., col = "red")
lines(df$Count, df$summary.90., col = "red")
lines(df$Count, df$other_mod, col = "black", type = "b", lwd = 1)
lines(df$Count, df$other_10, col = 'black')
lines(df$Count, df$other_90, col = "black")
lines(df$Count, df$true_injured, col = "green", type = "b", lwd = 1)
legend("topright", legend = c("Hierarchical: 10 pct", "Hierarchical: Mean Prediction", "Hierarchical: 90 pct", "True Injured", "Original Model"),
       col = c("red", "blue", "red", "green", "black"), lty = 1:2, cex = 0.8)
title(main = "Injury Prediction: Boston")


# without intervals 
plot(df$Count, df$summary.mean, type = "b", col = "blue", ylim = c(75,300), lty = 1, lwd = 1, 
      xlab = "Months",
      ylab = "Traffic Injuries")
lines(df$Count, df$other_mod, col = "black", type = "b", lwd = 1)
lines(df$Count, df$true_injured, col = "green", type = "b", lwd = 1)
legend("topright", legend = c("Original Model", "Hierarchical Prediction", "True Injured"),
       col = c("black", "blue", "green"), lty = 1:2, cex = 0.8)
title(main = "Injury Prediction: Boston (No CIs)")


# means absolute error before 
mae <- mean(abs(df$summary.mean[1:10] - df$true_injured[1:10]))
mae

#After 
mae <- mean(abs(df$summary.mean[11:20] - df$true_injured[11:20]))
mae

```


# Prior Predictive checks

## Boston
```{r}

# isolate city data
df_boston1 <- merged_df %>%
  filter(CITY == "BOSTON")

# get priors
brms::get_prior(Total_Injuries ~ 0 + arma(p = 1, q = 1), data = df_boston1)

prior1 <- c(set_prior("normal(0,100)", class = "ar"),
           set_prior("normal(0,100)", class = "ma"),
           set_prior("student_t(3, 0, 29.7)", lb = 0, class = "sigma"))


boston_arma_prior_predict <- brm(Total_Injuries ~ 0 + arma(p = 1, q = 1), data = df_boston1,
                                 seed = 1234,
                                 prior = prior1,
                                 sample_prior = "only",
                                 chains = 4,
                                 iter = 2000,
                                 cores = getOption("mc.cores", 4))



# generate data set using posterior_predict function
ynew_si <- posterior_predict(boston_arma_prior_predict)
dim(ynew_si)


# residuals
ytildehat_i <- apply(ynew_si, 2, min)
res <- df_boston1$Total_Injuries - ytildehat_i

plot(res, 
     xlab = "Index",
     ylab = "Residual")
title(main = "Residual Analysis: Simulated Data vs. Actual Data")
abline(h=0)

```



